Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2023-02-13 analysis/paper/paper.qmd
BiocManager

Our results show positive correlation between attributes throughout the Late Paleolithic (Panel A of @fig-corr-analysis). Some relationships such as BL and ML have stronger correlation (i.e. darker-blue points) than others (i.e. lighter-blue or almost invisible points). To understand temporal patterns in the modes of cultural transmission, we grouped our assemblages into the three Korean Late Paleolithic chronological phases summarized in @tbl-korean-chronology. Among the three chronological phases, we excluded Phase 1 from our analysis because there are only two complete stemmed points from Yonghodong and Sachang that date to this phase.

The correlations between attributes became stronger over time as we can clearly see in the Panel D of @fig-corr-analysis. The median for Phase 2 is much lower than `r threshold_value_cor`, which @bettinger1999point interpreted as a threshold to evaluate transmission biases, while the Phase 3 median is close to the threshold. Based on this results, we assume that the earlier transmission for stemmed points was more likely the results of guided variation. On the other hand, Phase 3 could be more associated with indirect bias.

@@

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making.The four assemblages include three from the Suyanggae site (SYG) and one from the Yongsandong site (YS). There are more variation in correlation coefficient in Phase 2 while the median values for the two assemblages from Phase 3 are similar to each other and close to `r threshold_value_cor`. There is no clear pattern between assemblages. SYG1_2 is from Phase 2 but it has the higher median than the assemblage from the same site in the later phase.

>>>>

Our results show positive correlations between attributes throughout the Late Paleolithic (Panel A of @fig-corr-analysis). Some relationships such as BL and ML have stronger correlation (i.e. darker-blue points) than others (i.e. lighter-blue or almost invisible points). To understand temporal patterns in the modes of cultural transmission, we grouped our assemblages into the three Korean Late Paleolithic chronological phases summarized in @tbl-korean-chronology. Among the three chronological phases, we excluded Phase 1 from our analysis because there are only two complete stemmed points from Yonghodong and Sachang that date to this phase.

Panel D of @fig-corr-analysis shows that the correlations between attributes became stronger over time. The median correlation value for Phase 2 is much lower than `r threshold_value_cor`, which @bettinger1999point interpreted as a threshold to distinguish between transmission biases, while the Phase 3 median is close to this threshold. value. <!-- this next sentence is probably best in the discussion section --> Based on this results, we assume that the earlier transmission for stemmed points was more likely the results of guided variation. On the other hand, Phase 3 could be more associated with indirect bias.

@@

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making. The four assemblages include three from Suyanggae (SYG) and one from Yongsandong (YS). There are more variation in correlation coefficient in Phase 2 while the median values for the two assemblages from Phase 3 are similar to each other and close to `r threshold_value_cor`. There is no clear pattern between assemblages. SYG1_2 is from Phase 2 but it has the higher median than the assemblage from the same site in the later phase.
